Toward \earlywarning\ detection of gravitational waves from compact binary coalescence

### 2012-01-10 article.tex
f_s -> f^0
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1459 ccdd05d5-56c1-461a-bf98-85f6876523b4

Searches for inspiral signals typically employ matched filter
banks that discretely sample the possible intrinsic parameters~\citep{findchirppaper}.  Suppose that the observed data $x[k]$ consists of a known, nominal signal $s[k]$ and additive, zero-mean noise $n[k]$,
%
$$
	x[k] = s[k] + n[k].
$$
%
A matched filter is a linear filter, defined as
%
$$
	y[k] = \sum_{n=0}^{N-1} h[n] \, x[k-n] = y_s[k] + y_n[k],
$$
%
where $y_s$ is the response of the filter to the signal alone and $y_n$ is the response of the signal to noise alone.  The matched filter's coefficients maximize the ratio of the expectation of the filter's instantaneous response to the variance in the filter's output:
%
$$
\textrm{signal to noise} = \frac{\expectation \left[ y[0] \right]^2}{\var \left[ y[k] \right]} = \frac{y_s[0]^2}{\var \left[ y_n[k] \right]}.
$$
%
It is well known~\citep[see, for example,][]{matched-filter} that if $n[k]$ is Gaussian and wide-sense stationary, then the optimum is obtained when
%
$$
\tilde{h}[n] = \tilde{s}^*[n] \, \tilde{S}^{-1}[n],
$$
%
up to an arbitrary multiplicative constant.  Here, $\tilde{h}[n]$, $\tilde{s}[n]$, and $\tilde{x}[n]$ are the discrete Fourier transforms (DFTs) of $h[k]$, $s[k]$, and $x[k]$, respectively; $\tilde{S}[n] = \expectation \left[ \tilde{n}[n] \tilde{n}^* [n] \right]$ is the folded, two-sided, discrete power spectrum of $n[k]$.  It is related to the continuous, one-sided power spectral density $S(f)$ through
$$
	\tilde{S}[n] =
	\begin{cases}
		S(n f_s / 2 N) / 2 & \textrm{if } 0 \leqslant n \leqslant N / 2 \\
		\tilde{S}[N - n] & \textrm{otherwise},
	\end{cases}
$$
where $N$ is the length of the filter and $f_s$ is the sample rate.  (In order to satisfy the Nyquist-Shannon sampling criterion, it is assumed that the detector's continuous noise power spectral density $S(f)$ vanishes for all $f > f_s / 2$, or alternatively, that the data is low-pass filtered prior to matched filtering.)  The discrete Fourier transform of the output is
%
\begin{multline}
\label{eq:matched-filter-fd}
\tilde{y}[n] = \tilde{s}^*[n] \, \tilde{S}^{-1}[n] \, \tilde{x}[n] \\
\equiv \left(\tilde{S}^{-1/2}[n] \, \tilde{s}[n]\right)^* \left(\tilde{S}^{-1/2}[n] \, \tilde{x}[n] \right).
\end{multline}
%
The placement of parentheses in equation~(\ref{eq:matched-filter-fd}) emphasizes that the matched filter can be thought of as a cross-correlation of a whitened version of the data with a whitened version of the nominal signal.  In this paper, we shall not describe the exact process by which the detector's noise power spectrum is estimated and deconvolved from the data; for the remainder of this paper we shall define $x[k]$ as the \emph{whitened} data stream.  Correspondingly, from this point on we shall use $h[k]$ to describe the \emph{whitened} templates, being the inverse DFT of $\left(\tilde{S}^{-1/2}[n] \, \tilde{s}[n]\right)^*$.

@@

\begin{thm}
	\label{thm:nyquist}
	\citep[after][p. 518]{oppenheim1997signals}
	Let $x(t)$ be a band-limited signal with continuous Fourier transform $\tilde{x}(f)$ such that $\tilde{x}(f') = 0 \; \forall \; f' : |f'| < f_M$.  Then $x(t)$ is uniquely determined by its discrete samples $x(n/f_s)$, $n = 0, \pm 1, \pm 2, \dots$, if $f_M < f_s / 2$.
\end{thm}

>>>>

Searches for inspiral signals typically employ matched filter
banks that discretely sample the possible intrinsic parameters~\citep{findchirppaper}.  Suppose that the observed data $x[k]$ consists of a known, nominal signal $s[k]$ and additive, zero-mean noise $n[k]$,
%
$$
	x[k] = s[k] + n[k].
$$
%
A matched filter is a linear filter, defined as
%
$$
	y[k] = \sum_{n=0}^{N-1} h[n] \, x[k-n] = y_s[k] + y_n[k],
$$
%
where $y_s$ is the response of the filter to the signal alone and $y_n$ is the response of the signal to noise alone.  The matched filter's coefficients maximize the ratio of the expectation of the filter's instantaneous response to the variance in the filter's output:
%
$$
\textrm{signal to noise} = \frac{\expectation \left[ y[0] \right]^2}{\var \left[ y[k] \right]} = \frac{y_s[0]^2}{\var \left[ y_n[k] \right]}.
$$
%
It is well known~\citep[see, for example,][]{matched-filter} that if $n[k]$ is Gaussian and wide-sense stationary, then the optimum is obtained when
%
$$
\tilde{h}[n] = \tilde{s}^*[n] \, \tilde{S}^{-1}[n],
$$
%
up to an arbitrary multiplicative constant.  Here, $\tilde{h}[n]$, $\tilde{s}[n]$, and $\tilde{x}[n]$ are the discrete Fourier transforms (DFTs) of $h[k]$, $s[k]$, and $x[k]$, respectively; $\tilde{S}[n] = \expectation \left[ \tilde{n}[n] \tilde{n}^* [n] \right]$ is the folded, two-sided, discrete power spectrum of $n[k]$.  It is related to the continuous, one-sided power spectral density $S(f)$ through
$$
	\tilde{S}[n] =
	\begin{cases}
		S(n f^0 / 2 N) / 2 & \textrm{if } 0 \leqslant n \leqslant N / 2 \\
		\tilde{S}[N - n] & \textrm{otherwise},
	\end{cases}
$$
where $N$ is the length of the filter and $f^0$ is the sample rate.  (In order to satisfy the Nyquist-Shannon sampling criterion, it is assumed that the detector's continuous noise power spectral density $S(f)$ vanishes for all $f > f^0 / 2$, or alternatively, that the data is low-pass filtered prior to matched filtering.)  The discrete Fourier transform of the output is
%
\begin{multline}
\label{eq:matched-filter-fd}
\tilde{y}[n] = \tilde{s}^*[n] \, \tilde{S}^{-1}[n] \, \tilde{x}[n] \\
\equiv \left(\tilde{S}^{-1/2}[n] \, \tilde{s}[n]\right)^* \left(\tilde{S}^{-1/2}[n] \, \tilde{x}[n] \right).
\end{multline}
%
The placement of parentheses in equation~(\ref{eq:matched-filter-fd}) emphasizes that the matched filter can be thought of as a cross-correlation of a whitened version of the data with a whitened version of the nominal signal.  In this paper, we shall not describe the exact process by which the detector's noise power spectrum is estimated and deconvolved from the data; for the remainder of this paper we shall define $x[k]$ as the \emph{whitened} data stream.  Correspondingly, from this point on we shall use $h[k]$ to describe the \emph{whitened} templates, being the inverse DFT of $\left(\tilde{S}^{-1/2}[n] \, \tilde{s}[n]\right)^*$.

@@

\begin{thm}
	\label{thm:nyquist}
	\citep[after][p. 518]{oppenheim1997signals}
	Let $x(t)$ be a band-limited signal with continuous Fourier transform $\tilde{x}(f)$ such that $\tilde{x}(f') = 0 \; \forall \; f' : |f'| < f_M$.  Then $x(t)$ is uniquely determined by its discrete samples $x(n/f^0)$, $n = 0, \pm 1, \pm 2, \dots$, if $f_M < f^0 / 2$.
\end{thm}
